---
title: "Homework 4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(classdata)
library(tidyverse)
library(ggplot2)
```

Loading in the Data
```{r}
choco = read.csv("https://ds202-at-isu.github.io/labs/data/choco.csv")
```

## Question Set 1

1. What is the overall number of chocolate bars rated? <br>
**Answer: 1852 number of chocolate bars rated**
```{r}
str(choco)
```
2. How does the number of ratings depend on the year? Draw a bar chart of the number of reports <br>
**Answer:  **
```{r}
choco %>% ggplot(aes(x = as.factor(Review.Date))) + 
  geom_bar() +
  ggtitle("") +
  xlab("Year")
```

## Question Set 2

1. How are ratings distributed? Draw a histogram of ratings and describe it. Don’t forget to mention outliers, if there are any. <br>
**Answer:  **

2. Do ratings depend on the cocoa percentage of a chocolate bar? <br>
**Answer:  **

3. How do ratings compare across different company locations? Focus on the three locations with the most ratings: <br>
**Answer:  **

## Your Own Question

1.  <br>
**Answer: **
