---
title: "Avengers Data"
author: "Polina Semenova"
date: "3/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```

Group Members: Michael Moore, Kateryna Mykhaylova<br>
Git Repo: [link](https://github.com/psemenova/DS202/tree/main/Homework/Avergers%20(practice4))

## Reading the data
```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
```


## Part One: Data tidying
The avenger data consists of ten columns for each avenger called Death[1-5] to Return[1-5]<br>

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the result `deaths`. <br>

Similarly, deal with the returns of characters.<br>

```{r}
av %>% mutate(Time = )
```


## Part two: Validation
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

**Polina Semenova**
```{r}

```

**Michael Moore**
```{r}

```

**Kateryna Mykhaylova**
```{r}

```
