---
title: "Homework 9 - Avengers Data"
author: "Polina Semenova"
date: "3/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```

Group Members: Michael Moore, Kateryna Mykhaylova<br>
Git Repo: [link](https://github.com/psemenova/DS202/tree/main/Homework/Homework%209)

## Reading the data
```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
```


## Part One: Data tidying
The avenger data consists of ten columns for each avenger called Death[1-5] to Return[1-5]<br>

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the result `deaths`. <br>

Similarly, deal with the returns of characters.<br>

```{r}
names(av) # columns

av0 = av %>% filter(Name.Alias != "")
av0 = av0 %>% filter(Name.Alias != "Vance Astrovik")

av1 = av0 %>% gather(key = time, value = death, c(11, 13, 15, 17, 19)) %>%
  select(Name.Alias, time, death) %>% 
  mutate(time = parse_number(time))

av2 = av0 %>% gather(key = time, value = return, c(12, 14, 16, 18, 20)) %>%
  select(Name.Alias, time, return) %>%
  mutate(time = parse_number(time))

av.neat = left_join(av1, av2, by = c("Name.Alias", "time"))
```


## Part two: Validation
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

**Polina Semenova**
```{r}

```

**Michael Moore**
```{r}

```

**Kateryna Mykhaylova**
```{r}

```
