---
title: "Homework 9 - Avengers Data"
author: "Polina Semenova"
date: "3/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```

Group Members: Michael Moore, Kateryna Mykhaylova<br>
Git Repo: [link](https://github.com/psemenova/DS202/tree/main/Homework/Homework%209)

## Reading the data
```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
```


## Part One: Data tidying
The avenger data consists of ten columns for each avenger called Death[1-5] to Return[1-5]<br>

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the result `deaths`. <br>

Similarly, deal with the returns of characters.<br>

```{r}
names(av) # columns

av0 = av %>% filter(Name.Alias != "")
av0 = av0 %>% filter(Name.Alias != "Vance Astrovik")

av1 = av0 %>% gather(key = time, value = death, c(11, 13, 15, 17, 19)) %>%
  select(Name.Alias, time, death) %>% 
  mutate(time = parse_number(time))

av2 = av0 %>% gather(key = time, value = return, c(12, 14, 16, 18, 20)) %>%
  select(Name.Alias, time, return) %>%
  mutate(time = parse_number(time))

av.neat = left_join(av1, av2, by = c("Name.Alias", "time"))
```


## Part two: Validation
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### Polina Semenova
**Statement:** Iron Man, Hulk, Captain America, Thor, Hawkeye, Black Widow, Scarlet Witch, Quicksilver and The Vision — every single one of them has died at least once in the course of their time Avenging in the comics. <br>
**Answer:** The table below shows all nine avengers show up at once. 
```{r}
avengers = av.neat %>% filter(Name.Alias %in% c("Anthony Edward \"Tony\" Stark", "Steven Rogers", "Clinton Francis Barton", "Thor Odinson", "Natalia Alianovna Romanova", "Wanda Maximoff", "Pietro Maximoff", "Victor Shade (alias)", "Robert Bruce Banner") & death == "YES")
head(avengers, 9)
```

### Michael Moore
**Statement:** In fact, Hawkeye died twice! <br>
**Answer:** This is proven true. There are only two rows in the data where Hawkeye (Clinton Francis Barton) died.
```{r}
hawkeye = av.neat %>% dplyr::filter(Name.Alias == "Clinton Francis Barton" & death == "YES")
nrow(hawkeye)
```

### Kateryna Mykhaylova
**Statement:** and on 57 occasions the individual made a comeback.<br>
**Answer:** This statement was wrong. As the data shows below, there were only 55 occasions in which the individuals made a comeback. 
```{r}
back = av.neat %>% filter(return == "YES") %>% summarise(n = n())
head(back)
```
